\papertitle

### 2020-01-28 paper/paper.tex
Software stack

The relevant software stack involved in executing workflows is depicted in \Cref{XXX}
\jk{insert picture about software stack here}
To solve a computational problem, we can consider describing the necessary steps from the start to the scientific insight generated by the workflow.
Cylc executes a workflow description comprising of the individual data processing steps and the dependencies among them.

@@

@@

So far, Cylc allowed specifying tasks and the dependencies among them.
We aim to enrich this information with characteristics for input, and the products. {\color{cyan}{What?!}}
An example workflow containing data products is illustrated in \Cref{fig:workflow}.
\sout{{\color{blue}{Arrows indicate dependencies between nodes that represent tasks and data.}}}
{\color{cyan}{Nodes represent tasks and data and arrows indicate dependencies among tasks.}}
In the example, Task\,1 needs two datasets to perform its work, it directly communicates with Task\,2 and produces Product\,1.
Most of the workflow can run automatically, except for the manual quality control of the products and the final data usage of Product\,3.
This last step represents the typical uncertainty of data reuse, i.e., it is unclear how Product\,3 will be used further.
In this approach, each task is annotated with the required inputs and the generated products.
Each data product is annotated with metadata such as data life cycle information, e.g., the value of data and how long should it be kept.

@@

File & \texttt{input.ds} & \texttt{tasks.ds} & \texttt{output.ds} \\ \hline

>>>>

@@

The relevant software stack involved in executing a workflow is depicted in \ref{fig:stages}.
As a preliminary, the user had specified a Cylc suite representing the workflow and provided scripts files for the individual tasks to execute; now the user enacts Cylc to start the workflow.
In the following, each stage of the execution is described further.

\begin{figure}[H]
  \centering
  \includegraphics[scale=1.4]{stages}
  \caption{Software Stack and Stages of Execution}
  \label{fig:stages}
\end{figure}

\begin{enumerate}
  \item \textbf{Cylc}\ \ Cylc parses the suite file, generate dependencies and defines a schedule for the execution.
  Cylc monitors the progress of the workflow.
  Once a task can be executed (dependencies are fulfilled), Cylc generates a Slurm script on the fly with the required metadata for Slurm that will invoke the user-defined script for that task, and runs Slurm to queue the job.

  \item \textbf{Slurm}\ \ Schedules the execution of the job considering the scheduling policy of the data center.
  Once the job is ready to be dispatched, i.e., resources are available and the job priority is the highest, it is started on the supercomputer.

  \item \textbf{Job}\ \ The user-provided script is executed on a single node assigned to the job.
  The script has set environment variables containing information about the Cylc execution (e.g., task in the workflow) and Slurm job (e.g., number of nodes).

  \item \textbf{Script}\ \ The shell-script executes the sequence of commands including applications that are parallel programs.
  Typically, to create a filename, the script calls Cylc commands with a template provided by the user and feeds this information to the application.
  This ultimately defines the storage location to be used.
  A script typically consist of multiple fine-grained commands and may even call multiple parallel applications.

  \item \textbf{Application}\ \ The user-specified parallel application runs taking the filenames as defined in a configuration file or as arguments.
  The application may use NetCDF or XIOS to perform the I/O.
\end{enumerate}

@@

So far, Cylc allowed specifying tasks and the dependencies among them.
We aim to enrich this information with characteristics for input, and output, i.e., the datasets used.
An example workflow containing input data sets and (intermediate) products is illustrated in \Cref{fig:workflow}.
Nodes represent tasks or data; arrows indicate dependencies among between them.
In the example, Task\,1 needs two datasets to perform its work, it directly communicates with Task\,2 and produces Product\,1.
Most of the workflow can run automatically, except for the manual quality control of the products and the final data usage of Product\,3.
This last step represents the typical uncertainty of data reuse, i.e., it is unclear how Product\,3 will be used further.
In this approach, each task is annotated with the required inputs and the generated products.
Each data product is annotated with metadata such as data life cycle information, e.g., the value of data and how long should it be kept.

@@

\jk{?} File & \texttt{input.ds} & \texttt{tasks.ds} & \texttt{output.ds} \\ \hline
